### 2022-06-16 tex/supplement.tex
Add softmax simplex derivation

>>>>

\subsection{Simplex softmax parameterization}

We define the transformation $\phi: \mathbb{R}^{n-1} \to \Delta^{n}: z \mapsto x_-$, where $x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix}$, $x_i = \frac{1}{r} e^{z_i}$ for $i \in \{1, \ldots, n-1\}$, and $r = 1 + \sum_{i=1}^{n-1} e^{z_i}$.

First we compute the scalar derivatives:
$$
\begin{aligned}
    \frac{\mathrm{d} r}{\mathrm{d} z_j} &= e^{z_j} = r x_j\\
    \frac{\mathrm{d} x_i}{\mathrm{d} z_j} &= \delta_{ij} \frac{1}{r} e^{z_i} - \frac{1}{r^2} e^{z_i} \frac{\mathrm{d} r}{\mathrm{d} z_j} = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
\end{aligned}
$$

where $\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0 &\text{otherwise}\end{cases}$ is the Kronecker delta.

If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the elements of $x$, then the Jacobian is

$$J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),$$

where $\boldsymbol{1}_n$ is the $n$-vector of ones.

Using Sylvester's determinant theorem, $|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 - \boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so

$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} z_i\right) \left(1 + \sum_{i=1}^{n-1} e^{z_i}\right)^{-n}$$
